Paradigms in Physics 2.0

### 2017-06-29 paper.tex
fix gray area
cut horizontal figure

\section{Changes made}
As a result of this process, we have implemented a number of changes
to our physics major curriculum.  The resulting curriculum is outlined
in Fig.~\ref{fig:schedule}.  The changes consisted of introducing two
new sophomore-level courses (which may be taken in the junior year) to
better prepare our majors for the their junior year.  We eliminated \emph{Modern
Physics} and the \emph{Classical Mechanics Capstone} in favor of the two new
sophomore-level courses.  We eliminated the \emph{Math Methods} course in
favor of ``\emph{Math Bits}'' integrated into the \emph{Paradigm} courses.
Finally, we restructured our nine 3-week \emph{Paradigm} courses into six
5-week \emph{Paradigm} courses.  We changed our computational physics requirement,
and reduced the number of electronics courses.

@@

\subsection{Math bits}
\davidsays{Discuss extra week and previous experience earlier.}
We chose to eliminate the \emph{Math Methods} course in favor of formal
\emph{Math Bits} in the \emph{Paradigm} courses.  Our previous \emph{Math Methods} course
posed a challenge regarding placement in the curriculum: it needed to
be sufficiently late for students to be adequately prepared
for advanced mathematical content, but in order to be functional in
the curriculum it should teach content that is required by other
physics courses, and thus should be earlier in the curriculum.  In
response to these issues, and in recognition that the majority of our
majors are not bound for graduate school, and have no need for
advanced math methods, we chose to integrate the needed math content
into the \emph{Paradigm} courses, and allow advanced students to take our
graduate-level \emph{Math Methods} course.  In order to ensure that our
students receive the math instruction they require, and based on
our previous experience with math-intensive weeks (the \davidsays{``extra''
weeks having 3-week courses fitted into a 10-week quarter})
in the \emph{Paradigm} courses, we chose to integrate one week of \emph{Math Bits} into
each \emph{Paradigm} course.  The \emph{Math Bits} for the entire year are treated as a
distinct teaching assignment, so students will be accustomed to the
treatment of math as a distinct subject supporting the physics
content.  This provides some relief from teaching for
professors teaching \emph{Paradigm} courses, and at the same time ensures
that they do not succumb to the temptation to short-change the math content
to the detriment of the students.
\corinnesays{This paragraph is awkward.  I'm still thinking about some rewording.  This might be a good paragraph to discuss this afternoon.}

@@

\subsubsection{Physics of Contemporary Challenges}
\corinnesays{You might want to use your paragraph structure here, rather than sub-sub sections.}

In \emph{The Physics of Contemporary Challenges} we prepare students to
apply physics concepts and physical reasoning skills to sustainable
energy issues, climate change mechanisms, space exploration and
puzzles in fundamental physics. These ``real-world'' topics are chosen
for either the societal need (energy and climate), and/or the human
need to explore (space and fundamental physics). By prioritizing
inclusion of the most engaging challenges, we aim to attract and
retain as many potential physics majors as possible.
\davidsays{Add in reference and note?}

@@

\subsubsection{Techniques of Theoretical Mechanics}

While the \emph{Challenges} course takes a more experimental/applied physics
focus, the \emph{Techniques of Theoretical Mechanics} course has a more
theoretical physics flavor. The theme of this course is the discussion
of strategies for making sense of physics problems and symbolic
problem solving. This sense-making includes coordinating and
interpreting symbolic expressions with conceptual understandings,
geometric relationships, and physical intuitions.

The \emph{Theoretical Mechanics} course is aimed at students who are taking
or have completed the last course in the introductory sequence. The
physics content of the course is advanced Newtonian mechanics,
introduction to Lagrangian and Hamiltonian techniques, and special
relativity. These topics are convenient
for explicit discussion of sense-making strategies in physics because
(a) these students have recently studied problems that serve as limiting
cases for these more complex problems, (b) students at this level are
transitioning from solving problems primarily involving numbers to
problems with only symbolic parameters, and (c) sense-making as strategies 
can be discussed for approaching a problem, evaluating
answers, and refining intuitions about relativistic and other
unfamiliar situations. For example, we teach students how to use
spacetime diagrams to develop a story about a relativistic situation
and how to use hyperbola trigonometry on these diagrams to perform Lorentz
transformations that can be checked against the results of algebraic
Lorentz transformations.  The sense-making goals of the course are on
an equal footing with the physics goals, and sense-making is integrated
explicitly in all aspects of the course: exams, homework and during
in-class activities. \davidsays{CITE HERE?} Our aim is that students will develop
sense-making skills that will improve their learning in advanced
courses and will be valued by their upper-division course instructors,
research advisors, and future employers.

@@

\section{Summary}
We have developed a significant change to the Paradigms
curriculum.  This change focuses on the first two big ideas underlying
the Paradigms project: (1) close attention to content ordering, and (2)
developing a consensus and understanding of the curriculum among our
faculty members.  Our commitment to using active engagement in our
upper-division curriculum remains unchanged.  While the specific
content ordering we developed may be interesting, the \emph{process} by
which we reached it and developed a faculty consensus is at the
essence of what makes the Paradigms in Physics program distinctive.
Emily van Zee, a researcher with experience in ..., 
\corinnesays{Emily, please add a short description of yourself} 
has completed a retrospective study of the process we used to develop the 
original Paradigms program and we have used her results extensively in 
designing and implementing the Paradigms~2.0 process.  She has continued 
to observe, study, and document our work as we have begun implementing 
the revised curriculum. 
\corinnesays{David:  Add a citation to http://physics.oregonstate.edu/portfolioswiki/toolkit:start}
The process of developing faculty understanding of the Paradigms is
ongoing, as we are having professors new to the department shadow
experienced faculty as they teach \emph{Paradigm} course, prior to teaching
the same course themselves.

@@

%% \begin{table*}[htbp]
%% \caption{Hypothetical student schedule for a student entering without
%%   calculus.  Students with AP Calculus can take the junior-year
%%   courses one year earlier, and students who start calculus-based
%%   physics in the Fall of their second year are not
%%   delayed.\label{schedule}}
%% \begin{ruledtabular}
%% \begin{tabular}{rlll}
%%   \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
%%   \hline
%%   1 & \mathcourse{Differential calculus}{251} &
%%   \mathcourse{Integral calculus}{252} &
%%   \mathcourse{Multivariable calculus}{254} \\
%%     & & \onecredit{Freshman Seminar} & \fourcredit{Calc Physics I} \\
%% \hline  2 & \mathcourse{Vector calculus}{255}
%%   & \mathcourse{Differential equations}{256}
%%   & \mathcourse{Linear algebra}{341} \\
%%   & \fourcredit{Calc Physics II} & \fourcredit{Calc Physics III} &\mathcourse{Series \&
%%     sequences}{253} \\
%%   && \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
%%   \\
%% \hline 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
%% \\
%%   & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
%% \\
%%   & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
%% \\
%%   & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
%% \\
%% \hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
%% \\
%% & \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
%% \\
%% & \onecredit{Research} & \onecredit{Research}
%% \end{tabular}
%% \end{ruledtabular}
%% \end{table*}

%% \begin{table*}[htbp]
%% \caption{Hypothetical transfer student schedule.  Most transfer
%%   students will have a heavier schedule in their transfer year due
%%   to need to catch up on required math courses.\label{schedule}}
%% \begin{ruledtabular}
%% \begin{tabular}{rlll}
%%   \textbf{Year} & \textbf{Fall} & \textbf{Winter} & \textbf{Spring} \\
%%   \hline
%% 3 & \paradigm{Quantum fundamentals} & \paradigm{Osciallations \& waves} & \paradigm{Static fields}
%% \\
%%   & \paradigm{Energy \& entropy} & \paradigm{Periodic systems} & \paradigm{Central forces}
%% \\
%%   & \onecredit{Computational lab I} & \onecredit{Computational lab II} & \onecredit{Computational lab III}
%% \\
%%   & \threecredit{Electronics} & \threecredit{Computer interfacing} & \onecredit{Research}
%% \\
%% & \mathcourse{Linear algebra}{341} & \noted{Challenges}{3} & \noted{Theoretical Mechanics}{3}
%% \\
%% &&\mathcourse{Vector calculus}{255}
%% \\
%% \hline 4 & \capstone{Electromagnetism capstone} & \capstone{Quantum capstone} & \capstone{Thermal capstone}
%% \\
%% & \onecredit{Thesis} & \onecredit{Thesis} & \onecredit{Thesis}
%% \\
%% & \onecredit{Research} & \onecredit{Research}
%% \end{tabular}
%% \end{ruledtabular}
%% \end{table*}

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{vertical-schedule}
\caption{Hypothetical student schedule in the new curriculum.
  Transfer students begin in year 3, and take the marked additional
  courses.\label{fig:schedule}}
\end{figure}
\section{Changes made}
As a result of this process, we have implemented a number of changes
to our physics major curriculum.  The resulting curriculum is outlined
in Fig.~\ref{fig:schedule}.  The changes consisted of introducing two
new sophomore-level courses (which may be taken in the junior year) to
better prepare our majors for the their junior year.  We eliminated \emph{Modern
Physics} and the \emph{Classical Mechanics Capstone} in favor of the two new
sophomore-level courses.  We eliminated the \emph{Math Methods} course in
favor of ``\emph{Math Bits}'' integrated into the \emph{Paradigm} courses.
Finally, we restructured our nine 3-week \emph{Paradigm} courses into six
5-week \emph{Paradigm} courses.  We changed our computational physics requirement,
and reduced the number of electronics courses.

@@

\subsection{Math bits}
Since Paradigms~1.0 was began, we introduced a
math-intensive week preceding two of the \emph{Paradigm} courses.  These math weeks provided
just-in-time preparation of the math skills required for those courses.
We have found these math weeks to be effective and popular with students.  In contrast to
these weeks the \emph{Math Methods} course is unwieldy, and challenging to place in
the curriculum at a time where it is helpful to students.  Moreover, some of the content
in \emph{Math Methods} was not actually required for our undergraduate courses, and only
needed for students bound for graduate school.  We chose therefore to eliminate the \emph{Math Methods} course in favor of \emph{Math Bits} consisting of a single week of 
just-in-time math content incorporated in each \emph{Paradigm} course.  Advanced
students may also take our graduate-level \emph{Math Methods} course.  
The \emph{Math Bits} for the entire year is taught by a single professor,
which provides continuity and coherence for the students.  
This provides some assistance for
professors teaching \emph{Paradigm} courses, and at the same time ensures
that they do not succumb to the temptation to short-change the math content
to the detriment of the students.

@@

\paragraph{Physics of Contemporary Challenges}
In this course we prepare students to
apply physics concepts and physical reasoning skills to sustainable
energy issues, climate change mechanisms, space exploration and
puzzles in fundamental physics. These ``real-world'' topics are chosen
for either the societal need (energy and climate), and/or the human
need to explore (space and fundamental physics). By prioritizing
inclusion of the most engaging challenges, we aim to attract and
retain as many potential physics majors as possible.
\davidsays{Add in reference and note?}

@@

\paragraph{Techniques of Theoretical Mechanics}

While the \emph{Challenges} course takes a more experimental/applied physics
focus, \emph{Theoretical Mechanics} has a more
theoretical physics flavor. The theme of this course is the discussion
of strategies for making sense of physics problems and symbolic
problem solving. This sense-making includes coordinating and
interpreting symbolic expressions with conceptual understandings,
geometric relationships, and physical intuitions.

This course is aimed at students who are taking
or have completed the last course in the introductory sequence. The
physics content of the course is advanced Newtonian mechanics,
introduction to Lagrangian and Hamiltonian techniques, and special
relativity. These topics are convenient
for explicit discussion of sense-making strategies in physics because
(a) these students have recently studied problems that serve as limiting
cases for these more complex problems, (b) students at this level are
transitioning from solving problems primarily involving numbers to
problems with only symbolic parameters, and (c) sense-making as strategies 
can be discussed for approaching a problem, evaluating
answers, and refining intuitions about relativistic and other
unfamiliar situations. For example, we teach students how to use
spacetime diagrams to develop a story about a relativistic situation
and how to use hyperbola trigonometry on these diagrams to perform Lorentz
transformations that can be checked against the results of algebraic
Lorentz transformations.  The sense-making goals of the course are on
an equal footing with the physics goals, and sense-making is integrated
explicitly in all aspects of the course: exams, homework and during
in-class activities. \davidsays{CITE HERE?} Our aim is that students will develop
sense-making skills that will improve their learning in advanced
courses and will be valued by their upper-division course instructors,
research advisors, and future employers.

@@

\section{Summary}
We have developed a significant change to the Paradigms
curriculum.  This change focuses on the first two big ideas underlying
the Paradigms project: (1) close attention to content ordering, and (2)
developing a consensus and understanding of the curriculum among our
faculty members.  Our commitment to using active engagement in our
upper-division curriculum remains unchanged.  While the specific
content ordering we developed may be interesting, the \emph{process} by
which we reached it and developed a faculty consensus is at the
essence of what makes the Paradigms in Physics program distinctive.
Emily van Zee, a researcher with experience in ..., 
\corinnesays{Emily, please add a short description of yourself} 
has completed a retrospective study of the process we used to develop the 
original Paradigms program and we have used her results extensively in 
designing and implementing the Paradigms~2.0 process.  She has continued 
to observe, study, and document our work as we have begun implementing 
the revised curriculum~\cite{lessonslearned}.
The process of developing faculty understanding of the Paradigms is
ongoing, as we are having professors new to the department shadow
experienced faculty as they teach \emph{Paradigm} course, prior to teaching
the same course themselves.

@@
